---
title: "p8105_hw1_cac2225"
author: "Courtney Chan"
date: "September 18, 2018"
output: github_document
---

#Problem 1

##Creating the dataframe

Writing a chunk of code to define the variables within the dataframe. 

First running the library package tidyverse so that the function tibble will be available.

Then naming the dataframe and starting the tibble function.

Seed is set to 1 so that results are reproducible.

Within the dataframe, "random_sample_ten" is the name of the random sample of 10 numeric values, taken from the uniform distribution of 0 to 5, using the runif function. 

Next the, the logical vector "greater_than_2" is defined based on the values that "random_samp_ten" will generate. Values within random_samp_ten that are greater than 2 will return a value of "True" within greater_than_2's column, while values within random_samp_ten that are equal to or less than 2 will return a value of "False".

A character vector, "random_words", that is length 10 is defined.

A factor vector, "color_cat", that comprises of ten different catagory labels is defined and turned into a factor using the factor function.

```{r Create dataframe}
library(tidyverse)

set.seed(1)

problem1_df = tibble(
  random_samp_ten = runif(10,0,5),
  greater_than_2 = random_samp_ten > 2,
  random_words = c("My", "name", "is", "Courtney", "Chan","and", "I", "love", "data", "science"),
  color_cat = factor(c("red","orange","yellow","green","blue","indigo","purple","black","white","grey" ))
)
```

Upon running this chunk of code, a table is generated in the environment with the four corresponding columns for each variable: ten random numeric values generated from the runif function, corresponding categorization of the ten random values as either greater than or equal to/less than 2 under a logical vector column, the 10 character vector and the 10 character factor vector.

##Taking the Means

The next chunk of code will try to take the mean of each variable in the dataframe.

```{r Take means}
mean(problem1_df$random_samp_ten)
mean(problem1_df$greater_than_2)
mean(problem1_df$random_words)
mean(problem1_df$color_cat)
```

The mean function can be applied to the random_samp_ten variable and to the greater_than_2 variable. The random_samp_ten variable is numeric and contains a list of numbers and thus the mean of these numbers can be calculated. greater_than_2 is a logical vector, that can be coded numerically, thus each "True" becomes a 1 and each "False" becomes a 0. The mean can then be taken and can be interpreted as a proportion of 1's within the list.

The mean function cannot be applied to the random_words variable because it is a character variable - it is comprised of letters and does not have a numeric meaning. Nor can it be applied to the color_cat variable because it is a factor variable, which are seen as catagories. Neither of them are numeric or logical variables - it would not make mathematical sense to try and calculate a mean value using non-numeric values.

##Applying the as.numeric function

```{r Apply as.numeric, eval = FALSE}
as.numeric(problem1_df$greater_than_2)
as.numeric(problem1_df$random_words)
as.numeric(problem1_df$color_cat)
```
As per instructions, the chunk is shown but the output is not, using eval = FALSE.

Upon applying the as.numeric function to the logical vector greater_than_2, instead of being recorded as "true" and "false," they're recoded as 1s and 0s. When the as.numeric function is applied, the character vector random_Words values, are recoded as NA. Applying as.numeric to the factor vector color_cat, causes the values to be recorded as numbers based on their alphabetical order. The as.numeric function tries to coerce variables into being defined as numeric.

```{r Converting character and factor to numeric}
as.numeric(as.factor(problem1_df$random_words))
as.numeric(as.character(problem1_df$color_cat))
```

After the catagory variable, random_Words, has been converted to a factor variable, and then numeric, its conversion follows the previous code's pattern. The random_words take on an alphabetical order as a factor vector and are numbered as such upon being converted to a numeric vector.

The factor vector color_cat, becomes a character vector, which upon being converted into a numeric vector can only display NA. There are no numerical values that can be associated with it.

#Problem 2

##Creating Dataframe 2

The package tidyverse is first loaded so that the tibble function can be used. The seed is set to 2 to allow reproducible results when the random samples of x and y are generated.

Rnorm function is used in this case, to generate random samples from a normal distribution.

```{r Create dataframe2}
library(tidyverse)

set.seed(2)

problem2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  sum_xandy = x + y > 0,
  vnum_sum = as.numeric(sum_xandy),
  vfac_sum = as.factor(sum_xandy)
)
```

##Short Description of dataframe problem2_df

The size of my dataset is `r dim(problem2_df)`.
The mean of x in my dataset is `r mean(problem2_df$x)`.
The median of x in my dataset is `r median(problem2_df$x)`.
The proportion of cases for which the logical vector is "True" is `r mean(problem2_df$sum_xandy)`.

##Creating scatterplot y vs. x

Code for scatterplot with coloring based on the logical variable, sum_xandy. False values are designated as red, while true values are designated as blue.

```{r Plot 1 color via sum_xandy}
plot_sum_xandy = ggplot(problem2_df, aes(color = sum_xandy, x = x, y = y)) + geom_point()
plot_sum_xandy
```

Code for scatterplot with coloring based on the numerical variable, vnum_sum. Values of 1s are designated as light blue and values of 0s are designated as dark blue.

```{r Plot 2 color via vnum_sum}
plot_vnum_sum = ggplot(problem2_df, aes(color = vnum_sum, x = x, y = y)) + geom_point()
plot_vnum_sum
```

Code for scatterplot with coloring based on the factor variable, vfac_sum. False values are designated as red and True is designated as blue.

```{r Plot 3 color via vfac_sum}
plot_vfac_sum = ggplot(problem2_df, aes(color = vfac_sum, x = x, y = y)) + geom_point()
plot_vfac_sum
```
